Omics-based Single-step Trait Predition

### 2017-03-15 reports/index.Rmd
Fix typos in Methods

---
title: Omics-based Single-step Trait Predition
author:
  - name: Matthias Westhues
    email: matthias.westhues@uni-hohenheim.de
    affiliation: University of Hohenheim
  - name: Claas Heuer
    email: cheuer@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Georg Thaller
    email: thaller@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Rohan Fernando
    email: rohan@iastate.edu
    affiliation: Iowa State University
  - name: Albrecht E. Melchinger
    email: melchinger@uni-hohenheim.de
    affiliation: University of Hohenheim
    footnote: Corresponding Author
address:
  - code: University of Hohenheim
    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
  - code: University of Kiel
    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
  - code: Iowa State University
    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: ["~/Dropbox/Paper/library.bib"]
abstract: |
  This is the abstract.

  It consists of two paragraphs.
---

>>>>

---
title: Omics-based Single-step Trait Predition
author:
  - name: Matthias Westhues
    email: matthias.westhues@uni-hohenheim.de
    affiliation: University of Hohenheim
  - name: Claas Heuer
    email: cheuer@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Georg Thaller
    email: thaller@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Rohan Fernando
    email: rohan@iastate.edu
    affiliation: Iowa State University
  - name: Albrecht E. Melchinger
    email: melchinger@uni-hohenheim.de
    affiliation: University of Hohenheim
    footnote: Corresponding Author
address:
  - code: University of Hohenheim
    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
  - code: University of Kiel
    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
  - code: Iowa State University
    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: ["~/Dropbox/Paper/library.bib"]
abstract: |
  $\dots$
---

### 2017-03-15 reports/methods.Rmd
Fix typos in Methods

### Hybrid data
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
SNP BeadChip MaizeSNP50 [@Ganal2011].
After filtering for minor allele frequency ($\geq 5$%), heterozygosity rate
($\geq 5$%) and call frequency ($\geq 95$%), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
After applying established normalization procedures [@Smyth2003;@Ritchie2007], 
gene expression BLUEs were computed for 1,323 transcripts in reference to 
established protocols [@Smyth2003;@Ritchie2007;@Frisch2010] using the R-package 
*limma* [@Ritchie2015a].

#### Diversity panel
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
[@Yang2014].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
we will evaluate six in our prediction models.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013;@Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes and was kept for further analyses [@Li2013].

## Population structure
Principal component analyses (PCA) were run on scaled and centered predictor 
matrices using the *pca()* function from the R-package *LEA* [@Frichot2015].
Estimates of ancestry coefficients were estimated based on parse nonnegative
matrix factorization algorithms [@Frichot2014] using the *snmf()* function 
in *LEA*.
Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
repetitions for each value of $K$. 
For each $K$, only the repetition with the lowest cross-entropy was kept for 
further analyses.

## Core Sampling
In order to evaluate the influence of the genetic constitution of the set of
genotypes that has only data on one out of two predictors, we generated core 
samples using the R-package *corehunter* [@Thachuk2009].
These core samples were based on the 149 maize inbred lines from the diversity
panel which had gene expression as well as genomic information.
The size of the core set was varied in increments of ten percentage points and 
ranged from 10% to 90% of all 149 inbred lines.
For assembling the various core sets, we first computed the Modifed Rogers (MR)
distance between each pair of genotypes and then aimed at maximizing the
average distance between each member of the selected core.
MR was computed using the *rogers.dist()*-function from the R-pacakge *poppr*
[@Kamvar2015].

@@

### Kernels
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effect have been accomodated by the dependent variable
(@Mrode2014, pp. 34, 311, 312).
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the $l$-th group of inbred lines and the 
$o$-th predictor ($\mathbf{W_{lo}}$) has dimensions $n \times p$ where $n$
pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
the number of features.
For the diversity panel maize lines, $n$ corresponds to the number of genotypes
whereas, in the case of the hybrid data, $n$ corresponds to the number of
parent lines in the corresponding heterotic group.
In generating kernels for each predictor, all features in $\mathbf{W_{lo}}$ were
centered and standardized to unit variance.
Then, each kernel can be defined as

\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{G}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
\end{equation}

where $W$ denotes the number of features [@VanRaden2008].
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{G}_{lo}$.

@@

\begin{equation} \label{eq:GBLUPModel}
  \mathbf{y} = \mu + 
  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
  \mathbf{\epsilon},
\end{equation}

where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
respectively, $\mu$ is the fixed model intercept, $\mathbf{Z}_{lo}$ is the 
corresponding design matrix associating the random genotype effects (*i.e.*, 
breeding values) of the lines in the $l$-th inbred line group and the $o$-th 
predictor type with $\mathbf{y}$.
Note that, in the case of hybrid data, $\mathbf{Z}_{lo}$ associates the random
general combining ability effects in the parental inbred lines (either Dent or
Flint, as expressed by $l$) with the vector of observed hybrid performance
$\mathbf{y}$.
The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
to $\mathbf{G}_{lo} \sigma^{2}_{{u}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
for the residual error.
All prediction models were implemented using the R-package *BGLR* [@Perez2014].

@@

The breeding values in a prediction model are in general given by:
\begin{equation} \label{eq:mrnaebv}
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

@@

Consider now the situation in which one predictor is complete in the sense that
it contains information on the whole population but a second predictor is
incomplete and covers only a subset of the population.
We can take a similar route as in @Fernando2014 and impute covariates of the 
incomplete predictor by using covariates of the complete predictor.
Let the subscript $1$ denote individuals covered only by the complete
predictor whereas individuals covered by both, the complete and the
incomplete predictor, are indicated by the subscript $2$.
Furhter, the covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align} \label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
[@Legarra2009].

@@

\begin{equation}
  (\#eq:single-step-model)
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Full} \boldsymbol{\epsilon}_{Full} + \mathbf{e},
\end{equation}

@@

\begin{equation}
 (\#eq:single-step-submatrices)
\mathbf{X} = 
\begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}_{Full} = 
\begin{bmatrix}
  Z_1 \\
  0 
 \end{bmatrix}
\end{equation}

@@

\begin{align}
 (\#eq:entangled-augmented-single-step-model)
\begin{bmatrix}
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

@@

\begin{equation}
 (\#eq:breeding-values)
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{Full} 
\end{equation}

@@

\begin{align}
 (\#eq:final-model)
\mathbf{y} &= \mathbf{Xb} +
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l{Full}} \boldsymbol{\epsilon}_{l{Full}} +
\mathbf{e}.
\end{align}

@@

### Model validation
For the validation of our predictions, we employed leave-one-out
cross-validation (LOOCV) routines.
In the case of the diversity panel maize inbred lines, LOOCV was done by simply 
using a single genotype as a hold-out sample, which will subsequently be 
predicted, while using all other inbred lines for model training.
This process is repeated until all inbred lines have been used once for testing
and $n - 1$ times for model training.

>>>>

### Hybrid data
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
BeadChip MaizeSNP50 [@Ganal2011].
After filtering for minor allele frequency ($\geq 5$%), heterozygosity rate
($\leq 5$%) and call frequency ($\geq 95$%), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
After applying established normalization procedures [@Smyth2003;@Ritchie2007], 
gene expression BLUEs were computed for 1,323 transcripts in reference to 
established protocols [@Smyth2003;@Ritchie2007;@Frisch2010] using the R-package 
*limma* [@Ritchie2015a].

### Diversity panel
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18$ to $30^{\circ}$N and from $102$ to $110^{\circ}$E [@Yang2014].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
six were used for this study.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013;@Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes and was kept for further analyses [@Li2013].

## Population structure
Principal component analyses (PCA) were run on scaled and centered predictor 
matrices using the *pca()* function from the R-package *LEA* [@Frichot2015].
Estimates of ancestry coefficients were estimated based on sparse nonnegative
matrix factorization algorithms [@Frichot2014] using the *snmf()* function 
in *LEA*.
Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
repetitions for each value of $K$. 
For each $K$, only the repetition with the lowest cross-entropy was kept for 
further analyses.

## Core Sampling
In order to evaluate the influence of the genetic constitution of the set of
genotypes that has only data on one out of two predictors, we generated core 
samples using the R-package *corehunter* [@Thachuk2009].
These core samples were based on the 149 maize inbred lines from the diversity
panel which had gene expression as well as genomic information.
The size of the core set was varied in increments of ten percentage points and 
ranged from 10% to 90% of all 149 inbred lines.
For assembling the various core sets, we first computed the Modifed Rogers (MR)
distance between each pair of genotypes and then selected a predefined subset
of genotypes that display maximum average distance between each other.
MR was computed using the *rogers.dist()* function from the R-pacakge *poppr*
[@Kamvar2015].

@@

### Kernels
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effects have been accomodated by the dependent variable [@Mrode2014].
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the $l$-th group of inbred lines and the 
$o$-th predictor ($\mathbf{W}_{lo}$) has dimensions $n \times p$ where $n$
pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
the number of features.
For the diversity panel maize lines, $n$ corresponds to the number of genotypes
whereas, in the case of the hybrid data, $n$ corresponds to the number of
parent lines in the corresponding heterotic group.
In generating kernels for each predictor, all features in $\mathbf{W}_{lo}$ were
centered and standardized to unit variance.
Then, each kernel can be defined as

\begin{equation} (\#eq:GenomicRelationship)
  \mathbf{K}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
\end{equation}

where $W_{lo}$ denotes the number of features [@VanRaden2008].
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{K}_{lo}$.

@@

\begin{equation} (\#eq:KBLUPModel)
  \mathbf{y} = \mu + 
  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
  \mathbf{\epsilon},
\end{equation}

where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
respectively, $\mu$ is the fixed model intercept, $\mathbf{Z}_{lo}$ is the 
corresponding design matrix associating the random genotype effects (*i.e.*, 
breeding values) of the lines in the $l$-th inbred line group and the $o$-th 
predictor type with $\mathbf{y}$.
Note that, in the case of hybrid data, $\mathbf{Z}_{lo}$ associates the random
general combining ability effects in the parental inbred lines (either Dent or
Flint, as expressed by $l$) with the vector of observed hybrid performance
$\mathbf{y}$.
The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
to $\mathbf{K}_{lo} \sigma^{2}_{{g}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
for the residual error.
All prediction models were implemented using the R-package *BGLR* [@Perez2014].

@@

The breeding values in a prediction model are in general given by:
\begin{equation} (\#eq:mrnaebv)
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

@@

Consider now the situation in which one predictor is complete in the sense that
it contains information on the whole population whereas a second predictor is
incomplete and covers only a subset of the population.
We can take a similar route as in @Fernando2014 and impute covariates of the 
incomplete predictor by using covariates of the complete predictor.
Let the subscript $1$ denote individuals covered only by the complete
predictor whereas individuals covered by both, the complete and the
incomplete predictor, are indicated by the subscript $2$.
Further, the covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align} (\#eq:mrna1)
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{K}_{11} - \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{K}_{21} = (\mathbf{K}^{11})^{-1}$
[@Legarra2009].

@@

\begin{equation}
  (\#eq:single-step-model)
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}^{Complete} \boldsymbol{\epsilon}^{Complete} + \mathbf{e},
\end{equation}

@@

\begin{equation}
 (\#eq:single-step-submatrices)
\mathbf{X} = 
\begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}^{Complete} = 
\begin{bmatrix}
  Z_1 \\
  0 
 \end{bmatrix}
\end{equation}

@@

\begin{align}
 (\#eq:entangled-augmented-single-step-model)
\begin{bmatrix}
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

@@

\begin{equation}
 (\#eq:breeding-values)
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}^{Complete} 
\end{equation}

@@

\begin{align}
 (\#eq:final-model)
\mathbf{y} &= \mathbf{Xb} +
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l}^{Complete} \boldsymbol{\epsilon}_{l}^{Complete} +
\mathbf{e}.
\end{align}

@@

### Model validation
For the validation of our predictions, we employed leave-one-out
cross-validation (LOOCV) routines.
In the case of the diversity panel maize inbred lines, LOOCV was performed by 
using a single genotype as a hold-out sample, which will subsequently be 
predicted, while using all other inbred lines for model training.
This process is repeated until all inbred lines have been used once for testing
and $n - 1$ times for model training.
